---
title: "ANLY 500 Final Project"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(mice)
library(dplyr)
library(RColorBrewer)
library(ggplot2)
library(reshape)
```

# Data Cleaning

- Unzip MERGED2014_15_PP.csv.zip, it is from College Scorecard webiste, not Kaggle. Since Kaggle do not have latest data.
- `mrc_table10` is data from Mobility Report Card data, they give each school a tier. Looks more useful than payscale school type. For detail, please see Codebook-MRC-Table-6.pdf.

## Read Data
```{r}
payscale_college_type = read.csv('salaries-by-college-type.csv')
payscale_college_type = payscale_college_type[, c(1,2)]
payscale_region = read.csv('salaries-by-region.csv')
payscale_region = payscale_region[, c(1,2)]

mrc_table10 = read.csv('mrc_table10.csv')
mrc_table10 = mrc_table10[, c(1, 12)]

# Treat "NULL" and "PrivacySuppressed" as NA when read
college_scorecard = read.csv('MERGED2014_15_PP.csv', na=c("NULL", "PrivacySuppressed"))
college_scorecard = college_scorecard[, c(3, 4,6,17,377,1639,1640,1642,1643,1645,1655,1656,1657,1661,1662,1665,1666,1668,1669,1671,1681,1682,1683,1687,1688,1697)]
```

## Merge Data

### Process School Name

Normalize all school name for merging.

```{r}
process_school_name = function(data) {
  data = sub(" \\(.*\\)", "", data)
  data = sub(" - ", "-", data)
  data = sub(", ", "-", data)
  data = sub("\\.", "", data)
  data = sub(" & ", " and ", data)
  data = sub("&", " and ", data)
  data = sub("St ", "Saint ", data)
  data
}

payscale_college_type$School.Name = process_school_name(payscale_college_type$School.Name)
nrow(payscale_college_type)

payscale_region$School.Name = process_school_name(payscale_region$School.Name)
nrow(payscale_region)

college_scorecard$INSTNM = process_school_name(college_scorecard$INSTNM)
nrow(college_scorecard)
```

### Merge Payscale data
```{r, results='hide'}
payscale = merge(payscale_college_type, payscale_region, by="School.Name", all = FALSE)

payscale %>% group_by(School.Name) %>% filter(n() > 1)

payscale_party = payscale %>%
  filter(School.Type == "Party")

payscale = payscale %>%
  filter(School.Type != "Party") %>%
  mutate(Is.Party = School.Name %in% payscale_party$School.Name)

nrow(payscale)
```

All the duplicate rows in Payscale data is because it duplicates all Party Schools. So we split whether or not is a party school into a separate column.

### Merge with College Scorecard
```{r, results='hide'}
scorecard_payscale = merge(payscale, college_scorecard, by.y = "INSTNM", by.x = "School.Name")

scorecard_payscale %>% group_by(School.Name) %>% filter(n() > 1)

scorecard_payscale = scorecard_payscale %>%
  filter( !((School.Name == 'Union College' & STABBR != 'NY') | 
            (School.Name == 'Wentworth Institute of Technology' & is.na(COSTT4_A))))

nrow(scorecard_payscale)
```

Two schools are duplicate after merged with College Scorecard data. After some search, we only keep Union College in New York because that is the only one in northwest. And we only keep one Wentworth Institute of Technology because the others do not have data.

### Merge with MRC
```{r, results='hide'}
data = merge(scorecard_payscale, mrc_table10, by.x = "OPEID6", by.y = "super_opeid")

data %>% group_by(School.Name) %>% filter(n() > 1)

nrow(data)
```

No duplicate in this step

### Other processing
```{r}
names(data)[names(data) == 'tier_name'] <- 'Tier'
earning_colnames = c("MN_EARN_WNE_P6", "MD_EARN_WNE_P6", "PCT25_EARN_WNE_P6", "PCT75_EARN_WNE_P6", "SD_EARN_WNE_P6", "MN_EARN_WNE_INC1_P6", "MN_EARN_WNE_INC2_P6", "MN_EARN_WNE_INC3_P6", "MN_EARN_WNE_MALE0_P6", "MN_EARN_WNE_MALE1_P6", "MD_EARN_WNE_P8", "MN_EARN_WNE_P10", "MD_EARN_WNE_P10", "PCT25_EARN_WNE_P10", "PCT75_EARN_WNE_P10", "SD_EARN_WNE_P10", "MN_EARN_WNE_INC1_P10", "MN_EARN_WNE_INC2_P10", "MN_EARN_WNE_INC3_P10", "MN_EARN_WNE_MALE0_P10", "MN_EARN_WNE_MALE1_P10")
cost_colnames = c("COSTT4_A")
attr_colnames = c("School.Name", "School.Type", "Region", "Is.Party", 
"STABBR", "CONTROL", "Tier")

data$CONTROL = factor(data$CONTROL, levels = c(1,2), labels = c("Public", "Private nonprofit"))
```

We reorder the column for easy inspection. And convert `CONTROL` into factor.

## Accuracy and Outlier
### Accurary & Missing value
```{r, results='hide', warning=FALSE}
summary(data)

percentmiss <- function(x){length(x[is.na(x)])/length(x)*100}

# process column first will get more records left
missing_col = apply(data, 2, percentmiss)
missing_col
delete <- which(missing_col > 5)
replace_col = data[,-delete]
dont_col = data[,delete]

missing_row = apply(replace_col, 1, percentmiss)
missing_row[missing_row > 5]
replace_row = subset(replace_col, missing_row <= 5)
dont_row = subset(replace_col, missing_row > 5)

# change to "cart" to avoid error, increase iteration to get reliable result
temp_no_miss = mice(replace_row, maxit=100, method='cart', seed=500)
no_miss = complete(temp_no_miss,1)

# combine data back
all_rows = rbind(dont_row, no_miss)
all_col = cbind(dont_col, all_rows)
```

There is no accuracy problem in the data. We use mice to complete the missing value for data meet 5% rule.

### Outlier
```{r}
# pass tolerance to prevent mahalanobis think it is singular matrix
attr_index = which(colnames(no_miss) %in% attr_colnames)
mahal <- mahalanobis(no_miss[-c(attr_index)],
                    colMeans(no_miss[-c(attr_index)],na.rm=TRUE),
                    cov(no_miss[-c(attr_index)], use = "pairwise.complete.obs"),
                    tol=1e-30)
cutoff = qchisq(1-.001,ncol(no_miss[-c(attr_index)]))
print(cutoff)
summary(mahal < cutoff)
noout = subset(no_miss, mahal < cutoff)

no_miss[mahal >= cutoff, c("School.Name", "COSTT4_A", "MN_EARN_WNE_P6", "MD_EARN_WNE_P6", "MN_EARN_WNE_P10", "MD_EARN_WNE_P10")]

data = all_col
```

The outliers are more than 5% of data, so we keep them.

## ROI Analysis
```{r}
## Break-even point calculation
data$BE_MN_P6 = data$COSTT4_A * 4 / data$MN_EARN_WNE_P6
data$BE_MD_P6 = data$COSTT4_A * 4 / data$MD_EARN_WNE_P6
data$BE_MN_P10 = data$COSTT4_A * 4 / data$MN_EARN_WNE_P10
data$BE_MD_P10 = data$COSTT4_A * 4 / data$MD_EARN_WNE_P10

## ROI USING EARNINGS of beigining salary
data$ROI_MN_P6 = (data$MN_EARN_WNE_P6 - data$COSTT4_A) / data$COSTT4_A * 100
data$ROI_MD_P6 = (data$MD_EARN_WNE_P6 - data$COSTT4_A) / data$COSTT4_A * 100

## ROI USING MEAN EARNINGS of 6-year salary
data$ROI_MN_P10 = (data$MN_EARN_WNE_P10 - data$COSTT4_A) / data$COSTT4_A * 100
data$ROI_MD_P10 = (data$MD_EARN_WNE_P10 - data$COSTT4_A) / data$COSTT4_A * 100
roi_colnames = c("BE_MN_P6", "BE_MD_P6", "BE_MN_P10", "BE_MD_P10", "ROI_MN_P6", "ROI_MD_P6", "ROI_MN_P10", "ROI_MD_P10")
```

### Data Output

```{r}
data_gender_p6 = melt(data[c("School.Name", "MN_EARN_WNE_MALE0_P6", "MN_EARN_WNE_MALE1_P6")], id=c("School.Name"), variable_name="Gender")
levels(data_gender_p6$Gender) = c("Male", "Female")
colnames(data_gender_p6)[3] = "MN_EARN_WNE_GENDER_P6"

data_gender_p10 = melt(data[c("School.Name", "MN_EARN_WNE_MALE0_P10", "MN_EARN_WNE_MALE1_P10")], id=c("School.Name"), variable_name="Gender")
levels(data_gender_p10$Gender) = c("Male", "Female")
colnames(data_gender_p10)[3] = "MN_EARN_WNE_GENDER_P10"

data_inc_p6 = melt(data[c("School.Name", "MN_EARN_WNE_INC1_P6", "MN_EARN_WNE_INC2_P6", "MN_EARN_WNE_INC3_P6")], id=c("School.Name"), variable_name="Family.Income.Tercile")
colnames(data_inc_p6)[3] = "MN_EARN_WNE_INC_P6"
levels(data_inc_p6$Family.Income.Tercile) = c("Lowest Income Tercile", "Middle Income Tercile", "Highest Income Tercile")

data_inc_p10 = melt(data[c("School.Name", "MN_EARN_WNE_INC1_P10", "MN_EARN_WNE_INC2_P10", "MN_EARN_WNE_INC3_P10")], id=c("School.Name"), variable_name="Family.Income.Tercile")
colnames(data_inc_p10)[3] = "MN_EARN_WNE_INC_P10"
levels(data_inc_p10$Family.Income.Tercile) = c("Lowest Income Tercile", "Middle Income Tercile", "Highest Income Tercile")

data = data[c(attr_colnames, cost_colnames, earning_colnames, roi_colnames)]
# write_csv(data, "data_cleaned.csv")
# write_csv(data_gender_p6, "data_gender_p6.csv")
# write_csv(data_gender_p10, "data_gender_p10.csv")
# write_csv(data_inc_p6, "data_inc_p6.csv")
# write_csv(data_inc_p10, "data_inc_p10.csv")
```

# Algorithm and Models
```{r}
##Read the 4th dataset
salary_degree <-
  read_csv(
    'degrees-that-pay-back.csv',
    col_names = c(
      "major",
      "start_med_slry",
      "mid_car_slry",
      "percent_chng",
      "mid_car_10th",
      "mid_car_25th",
      "mid_car_75th",
      "mid_car_90th"
    ),
    col_types = "cnndnnnn", # specify column types to coerce '$' to numeric
    skip = 1 # names specified, skip header
  ) 

#####
cleanup<-theme(panel.grid.major = element_blank(),
               panel.grid.minor = element_blank(),
               panel.background = element_blank(),
               axis.line.x = element_line(color = 'black'),
               axis.line.y = element_line(color = 'black'),
               legend.key = element_rect(fill = 'white'),
               text = element_text(size = 12)) 
##  Plot degree v/s starting salary
p1 <- ggplot(salary_degree, aes(x = reorder(major, start_med_slry), start_med_slry)) +
  geom_col(fill = "blue", alpha = 0.5) +
  geom_col(aes(x = reorder(major, mid_car_slry), mid_car_slry), alpha = 0.3) +
  geom_text(aes(label = (start_med_slry)), size = 3, hjust = 1.1, col="yellow")+
   xlab(NULL) +
  ylab("Salary in $")+
  coord_flip() +
  ggtitle("Starting salary v/s mid career salary in $")+
  cleanup
p1

## Top 10 school 

accent_colors_edit <- brewer.pal(n = 5, "Pastel1")[c(1:3, 5)]  # keep colors consistent for plot w/o 'party'

top10_colleges <- data %>%
  select(School.Name, School.Type, MD_EARN_WNE_P8) %>%
  arrange(desc(MD_EARN_WNE_P8)) %>%
  top_n(10)
top10_colleges
ggplot(top10_colleges, aes(reorder(School.Name, MD_EARN_WNE_P8), MD_EARN_WNE_P8, fill = School.Type)) +
  geom_col(alpha = 0.8) +
  scale_fill_manual(values = accent_colors_edit) +
  geom_text(aes(label =(MD_EARN_WNE_P8)), hjust = 1.1, color = 'gray30') +
   xlab(NULL) +
  ggtitle("Top 10 colleges")+
  ylab("Mid career salary in $")+
   coord_flip()
## Any correlation b/w starting salary and mid career salary

ggplot(data, aes(MD_EARN_WNE_P6, MD_EARN_WNE_P8)) +
  geom_point(alpha = 0.6) +
  geom_smooth(se = F) +
  xlab("Starting salary in $")+
  ylab("Mid career salary in $")
 
 
## Count by college type

ggplot(data, aes(School.Type)) +
  geom_bar(color = 'black', alpha = 0.8)
  
##  Region and type
  ggplot(data, aes(Region, fill = School.Type)) +
    geom_bar(position = 'dodge', alpha = 0.8, color = 'gray20') +
    scale_fill_brewer(palette = 'Pastel1') +
    theme(legend.position = "top")  

# How do starting salary and mid-career median salary differ over region and type? 

#Below is a look at the mean starting and mid-career salaries over these two categories.
  ggplot(data, aes(reorder(Region, MN_EARN_WNE_P6), MD_EARN_WNE_P8, fill = School.Type)) +
    stat_summary(geom = 'col', position = 'dodge', alpha = 0.6) +
    stat_summary(aes(Region, MN_EARN_WNE_P6, fill = School.Type),
                 geom = 'col', position = 'dodge') +
    scale_fill_brewer(palette = 'Pastel1') +
      xlab('Region') +
    ylab('Salaries in $') +
    ggtitle('Mean starting and Mid-career median salaries') +
    coord_flip()

```
